Faster generic IND-CCA2 secure KEM using ``encrypt-then-MAC''

### 2024-10-01 etm-transform/manuscript/main.tex
Started with rewriting introduction

%%%% 7. PAPER CONTENT %%%%
\section{Introduction}
The Fujisaki-Okamoto transformation \cite{DBLP:conf/crypto/FujisakiO99} is a generic construction that takes cryptographic primitives of lesser security and constructs a public-key encryption scheme with indistinguishability under adaptive chosen ciphertext attacks. Later works \cite{hofheinz2017modular}\cite{DBLP:conf/asiacrypt/HovelmannsHM22} extended the original transformation to the construction of key encapsulation mechanism, which has been adopted by many post-quantum schemes such as Kyber \cite{bos2018crystals} (later standardized into ML-KEM \cite{nist2024module}), FrodoKEM \cite{bos2016frodo}, and SABER \cite{d2018saber}.

The current state of the FO transformation enjoys proven tight security bound and quantum resistance, but also leaves many open problems. One such problem is the substantial computational cost of using \emph{de-randomization} and \emph{re-encryption} \cite{bernstein2018towards} for providing ciphertext integrity. In many post-quantum schemes, including ML-KEM, the input encryption routine is substantially more expensive than the input decryption routine, so running the encryption as a subroutine substantially increases the runtime cost of the decapsulation routine. While not the focus of this project, \emph{re-encryption} also introduces risks of side-channel vulnerabilities that may expose the plaintext or the secret key, as demonstrated in \cite{ravi2019generic} and \cite{ueno2022curse}.

We are inspired by how ciphertext integrity is achieved in symmetric cryptography: given a semantically secure symmetric cipher and an existentially unforgeable message authentication code, combining them using in a pattern called ``encrypt-then-MAC'' provides proven authenticated encryption \cite{bellare2000authenticated}. ``encrypt-then-MAC'' is now the most widely accepted method for doing symmetric encryption as AES-GCM \cite{mcgrew2004galois} and ChaCha20-Poly1305 \cite{nir2018chacha20}.

The main challenge in applying ``encrypt-then-MAC'' to public-key cryptography is the lack of a pre-shared symmetric key. Inspired by hybrid public-key encryption schemes such as \cite{DBLP:conf/crypto/FujisakiO99} and \cite{DBLP:conf/crypto/CramerS98}, we proposed to derive the symmetric key by hashing the plaintext message. In section \ref{sec:main-results}, we prove that under the random oracle model, if the input public-key encryption scheme is one-way secure against plaintext-checking attack and the input message authentication code is one-time existentially unforgeable, then the transformed key encapsulation mechanism is IND-CCA2 secure.

In section \ref{sec:implementation}, we instantiate concrete instances of our constructions using ML-KEM. Our experimental results showed that replacing re-encryption with computing authenticator leads to significant performance improvements in the decapsulation routine while incurring only minimal runtime overhead in the encapsulation routine and a small increase in ciphertext size.

>>>>

%%%% 7. PAPER CONTENT %%%%
\section{Introduction}
Key encapsulation mechanism (KEM) is a cryptographic primitive that allows two parties to establish a shared secret over an insecure channel. The most widely accepted security standard for a KEM is \emph{indistinguishability under adaptive chosen-ciphertext attack (IND-CCA2)}: no efficient adversary can distinguish a shared secret obtained by running encapsulation from random noise even with access to a decapsulation oracle throughout the attack. KEM is usually paired with a data encapsulation mechanism (DEM), such as AES-GCM or ChaCha20-Poly1305 when constructing secure communication protocol.

Provable IND-CCA2 security is difficult to achieve from scratch. One popular approach to constructing secure KEM is by starting with a public-key encryption (PKE) scheme with weaker security properties, then adding additional mechanisms for ensuring ciphertext integrity. One particularly successful example is the Fujisaki-Okamoto transformation, originally proposed in 1999 by Fujisaki Eiichiro and Okamoto Tatsuyaki in their seminal paper \cite{DBLP:conf/crypto/FujisakiO99}. The original Fujisaki-Okamoto transformation combines an OW-CPA secure PKE with an IND-CPA secure symmetric cipher into a hybrid PKE that achieves the full IND-CCA2 security in the random oracle model. Subsequent works improved on the original construction, culminating in the 2017 paper \cite{hofheinz2017modular} by Hofheinz, Hovelmann, and Kiltz, which provided a modular KEM construction with tight IND-CCA2 security reduction against classical adversary and proven (though non-tight) IND-CCA2 security reduction against quantum adversaries.

The modular Fujisaki-Okamoto KEM transformation is remarkably successful. It was adopted by many submissions to NIST's post-quantum cryptography competition, including Kyber \cite{bos2018crystals}\cite{avanzi2019crystals}, Saber \cite{d2018saber}, FrodoKEM \cite{bos2016frodo}, and classic McEliece \cite{} among others. When Kyber was standardized by NIST in FIPS 203 ``Module-lattice key-encapsulation mechanism'' (ML-KEM) \cite{FIPS203}, it kept the Fujisaki-Okamoto transformation in its KEM construction. However, the Fujisaki-Okamoto transformation is not perfect. Among its shortcomings is the use of \emph{de-randomization} and \emph{re-encryption} for ensuring ciphertext integrity: \begin{itemize}
    \item \emph{Computational inefficiency.}
    \item \emph{Side-channel vulnerability}
\end{itemize}
